---
title: "Project_Report"
author: "Joshua Austin"
output:
  bookdown::pdf_document2:
      toc: FALSE
bibliography: references.bib
nocite: '@*'
date: "5/11/2021"
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72

---

```{r setup, echo=FALSE, include=FALSE}
library(ThreeTests)
library(dplyr)

knitr::opts_chunk$set(echo = F, message=FALSE, comment =NA)



# Change the data with csv to report on a new dataset
  Reportdata <- ProjectData #read.csv("~Reportdataatreport.csv")


# To change parameters for any of the tests change filters below
  signif <- 0.05
  
  #details for a new linear model (question 1)
  LM_xdata <- Reportdata$height #change height to new column name
  LM_ydata <- Reportdata$weight
  LM_xname <- "height"
  LM_yname <- "weight"
  

  # change the t-test data
  tt_xdata <- (Reportdata %>% 
    filter(gender == "Male"))$height #change filter and height to suit new data
  tt_ydata <- (Reportdata %>% 
    filter(gender == "Female"))$height
  tt_cata <- "height"
  tt_xname <- "Male"
  tt_yname <- "Female"
  
  
  # chi-squared categorical test
  chi_xdata <- Reportdata$gender
  chi_ydata <- Reportdata$phys
  chi_xname <- "gender"
  chi_yname <- "physical activity"
  


```



# Abstract


An analysis of `r count(Reportdata)` observations in a dataset that includes the variables `r unique(c(LM_xname, LM_yname, tt_cata, chi_xname, chi_yname))`. The report looks at if there is a linear relationship between `r LM_xname` and `r LM_yname`, if the mean `r tt_cata` differs for `r tt_xname` and `r tt_yname`, and if there is a relationship between categorical variables `r chi_xname` and `r chi_yname`.


          
## Introduction
          

This report will work to answer three questions using a variety of different methods and statistical tests.

The first question is if a linear relationship existed between `r LM_xname` and `r LM_yname`. The relationship will be determined using *simple linear regression* and find, if it exists, the underlying *linear model*.

The next question is if the mean `r tt_cata` of `r tt_xname` and `r tt_yname` is the same. This is explored using a two sample *t-test* to compare the means of both groups.

The final question determines if there is an association between `r chi_xname` and `r chi_yname`. A *chi-squared* test is used to compare the spread of observation between the groups compared to what the expected observations would be if there is no association between the categories (@LockPattiFrazer2019S).

Each question has been address in a different way so have been separated in the following sections.


\newpage
# Methodology 

## Question 1

To test if there is a linear relationship between `r LM_xname` and `r LM_yname` a simple linear regression analysis is run. The linear regression will create a linear formula to relate the means of `r LM_xname` and `r LM_yname` before reviewing the residuals, difference between the observed and formulated values, to determine if the formula is reliable. Before the regression can be completed all assumptions need to be checked (@alma99244562776602171).

```{r lmhyp, echo=FALSE, fig.height=2.6, fig.width=6}
#The package includes a LinearModel function that uses the project dataset by default (the data was saved as a dataset included with the package under ProjectData)

LMfit <- LinearModel(LM_xdata, LM_ydata, LM_xname, LM_yname, signif)
LMfit$assumptions


LMfit$plot1

LMfit$plot2

LMfit$plot3


LMfit$hypothesis
```

The linear model fitted is;

\begin{equation}
`r LM_yname` = `r LMfit$lm_data$coefficients[2]` +  `r LMfit$lm_data$coefficients[1]` * `r LM_xname`\text{ + Epsilon.}
\end{equation}


## Question 2

A t-test will be used to check if there is a statistically significant difference between the mean `r tt_cata` of `r tt_xname` and `r tt_yname`. This test will compare the mean and variance of each group to determine to likelihood that there isn't actually a difference, and observed difference is due to random chance.



```{r ttestmeth, echo=FALSE, fig.height=4}
#The package includes a ttest_report.


ttestfit <-   ttest_report(tt_xdata, tt_ydata, tt_cata, tt_xname, tt_yname, signif)
ttestfit$assumptions
ttestfit$plot1
```
```{r ttestmeth2, echo=FALSE, fig.height=4}


ttestfit$plot2
ttestfit$hypothesis
```

  
The t-test will return a test statistic and corresponding p-value. The p-value invidacted the likelihood that the oberved difference in groups would occur if they can from the same distribution, a small p-value means that its unlikely that they came from the same distribution and there being a difference between the groups is likely.


## Question 3

To test if there is an association between categorical variables, `r chi_xname` and `r chi_yname` a *chi-squared* test is used. The test will compare the observed distribution between the groups with what would have been expected if there was infact no relationship.

```{r chimeth, echo=FALSE}
#The package includes a ChiSquared function.


Chifit <-  ChiSquared(chi_xdata, chi_ydata, chi_xname, chi_yname, signif)
Chifit$assumptions

Chifit$hypothesis
```

The Chi-Squared test will calculate a value for each cell using the below formula, all of the cells are then summed to find the test statistic and corresponding p-value.

\begin{equation}
\label{eqn:chisquared}
    \quad \frac{\left(\text{observed} - \text{expected}\right)^2}{\text{expected}}  
\end{equation}


\newpage
# Results

## Question 1

```{r}
LMfit$Report
LMfit$decision
```

    

## Question 2

The t-test gives the result;
```{r}
ttestfit$test.stats
```
Note the p-value listed towards the top may be rounded due to R limitations.

The test indicated to `r ttestfit$decision`.



## Question 3

The $\chi$-squared tests statistic and corresponding p-value are;
```{r}
Chifit$Chi.stats
Chifit$decision
```


\newpage
# Conclusion

After testing all 3 questions using a variety of statistical methods we conclude;


1. Is there a linear relationship between `r LM_xname` and `r LM_yname`?
  `r LMfit$conclusion`

2. Is the mean `r tt_cata` of `r tt_xname` and `r tt_yname` the same?
  `r ttestfit$conclusion`

3. Is there any association between `r chi_xname` and `r chi_yname`?
  `r Chifit$conclusion`
  
 
# References


  
